Carbonate chemistry at Ny-Ålesund

### 2023-03-09 awipev-CO2.Rmd
Added availability discrete data

library("RColorBrewer")
if (!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
if (!require("seacarb")) install.packages("seacarb")
library("seacarb")
if (!require("gridExtra")) install.packages("gridExtra")
library("gridExtra")
if (!require("lubridate")) install.packages("lubridate")
library("lubridate")
if (!require("grid")) install.packages("grid")
library(grid)
if (!require("viridis")) install.packages("viridis")
library("viridis")
if (!require("dygraphs")) install.packages("dygraphs")
library("dygraphs")
if (!require("knitr")) install.packages("knitr")
require("knitr")
if (!require("lmodel2")) install.packages("lmodel2")
library("lmodel2")
if (!require("captioner")) install.packages("captioner")
library("captioner")
if (!require("xts")) install.packages("xts")
library("xts")
if (!require("scales")) install.packages("scales")
library("scales")
if (!require("skimr")) install.packages("skimr")
library("skimr")
if (!require("FNN")) install.packages("FNN")
library("FNN")
if (!require("patchwork")) install.packages("patchwork")
library("patchwork")
if (!require("pander")) install.packages("pander")
library("pander")
if (!require("tsibble")) install.packages("tsibble")
library("tsibble")
if (!require("slider")) install.packages("slider")
library("slider")
if (!require("kableExtra")) install.packages("kableExtra")
library("kableExtra")
if (!require("readxl")) install.packages("readxl")
library("readxl")
if (!require("oce")) install.packages("oce")
library(oce)
if (!require("pangaear")) install.packages("pangaear")
library("pangaear")
if (!require("latex2exp")) install.packages("latex2exp")
library("latex2exp")
if (!require("gtsummary")) install.packages("gtsummary")
library("gtsummary")
if (!require("todor")) install.packages("todor")
library("todor")
if (!require("ggpmisc")) install.packages("ggpmisc")
library("ggpmisc")
if (!require("tidyselect")) install.packages("tidyselect")
library("tidyselect")
if (!require("usethis")) install.packages("usethis")
library("usethis")
if (!require("ggOceanMaps")) install.packages("ggOceanMaps")
library("ggOceanMaps")

@@

```{r, echo=FALSE, message= FALSE, warning=FALSE, out.width="80%",fig.show='hold', fig.align="center"}
d_min <- readRDS(file = paste0(path, "fb_awipev-co2_server/ny-alesund/data/NRT_data/all_nydata_minute.rds")) %>%
dplyr::select(datetime, c(pco2_corr_qf, sal_fb_qf, sal_insitu_ctd_qf, 
                          temp_fb_qf, temp_insitu_11m_qf, ph_dur_qf, phINT_qf))

s1 <- summary(as.factor(d_min$pco2_corr_qf))*100/length(d_min$pco2_corr_qf)
s2 <- summary(as.factor(d_min$sal_fb_qf))*100/length(d_min$sal_fb_qf)
s3 <- summary(as.factor(d_min$sal_insitu_ctd_qf))*100/length(d_min$sal_insitu_ctd_qf)
s4 <- summary(as.factor(d_min$temp_fb_qf))*100/length(d_min$temp_fb_qf)
s5 <- summary(as.factor(d_min$temp_insitu_11m_qf))*100/length(d_min$temp_insitu_11m_qf)
# durafet and seafet started later
x <- filter(d_min, datetime >= as.POSIXct("2017-08-29 10:00:00"))
#s6 <- summary(as.factor(x$temp_dur_qf))*100/length(x$temp_dur_qf)
s7 <- summary(as.factor(x$ph_dur_qf))*100/length(x$ph_dur_qf)
x <- filter(d_min, datetime >= as.POSIXct("2017-08-24 19:00:00"))
s8 <- summary(as.factor(x$phINT_qf))*100/length(x$phINT_qf)
z <- t(bind_rows(s1,s2,s3,s4,s5,s7,s8)) %>%
  as.data.frame() %>%
         rownames_to_column(var = "flag") %>% 
         as_tibble()
labels <- c("Good", "Not usable", "Outside regional range", "Outliers (despike)", 
            "Failed visual inspection", "Not deployed or operated", "Acid flush", "Failed visual inspection")
colnames(z) <- c("Flag", "pCO2", "Salinity FerryBox", "In situ salinity", 
                 "Temperature FerryBox", "In situ temperature", "Durafet pH", "seaFET pH")
p_avail <- pivot_longer(z, -Flag) %>% 
  drop_na() %>% 
  ggplot(aes(x = name, y = value, fill = Flag)) +
  geom_bar(position = position_stack(reverse = TRUE), stat="identity", width = 0.3) +
  scale_x_discrete(
    labels = c(
      "Flag",
      'pCO2' = parse(text = TeX('$pCO_2$')),
      "Salinity FerryBox",
      "In situ salinity",
      "Temperature FerryBox",
      "In situ temperature",
      "Durafet pH",
      "seaFET pH"
    )
  ) +
  coord_flip() +
  labs(x="", y = "Distribution of quality flags (%)") +
  Mytheme(size_labs = 14) +
  theme(legend.title = element_text(colour="black", size=14, face="bold"),
        axis.title.x = element_text(colour="black", size=14, face="bold")) #+
  #theme(plot.margin = margin(t = -8, r = 0.0, b = -8, l = 0.0, unit = "cm"))
#kable(z, digits = 1)
d2 <- d_all %>% 
  dplyr::select(montht, pco2_fb, phINT)
count_pco2_fb_no_NA <- sum(!is.na(d2$pco2_fb)) # total non-NA pCO2 values
count_phINT_no_NA <-  sum(!is.na(d2$phINT)) # total non-NA pH seafet values
# calculate the distribution of non-NA pCO2 and pH values per month
d2_summary <- d2 %>% 
  dplyr::group_by(montht) %>% 
  dplyr::summarise(pco2 = round(100 * sum(!is.na(pco2_fb))/1000, digits = 1),
                   pH = round(100 * sum(!is.na(phINT))/1000, digits = 1),
                   pco2_pc = round(100 * sum(!is.na(pco2_fb))/count_pco2_fb_no_NA, digits = 1),
                   pH_pc = round(100 * sum(!is.na(phINT))/count_phINT_no_NA, digits = 1)
                   )
# plot the distribution of non-NA pCO2 and pH values per month
p_avail_month <- d2_summary %>% 
  tidyr::pivot_longer(cols = c(pco2, pH)) %>% 
  ggplot() +
  geom_col(aes(x = montht, y = value, fill = name), position = "dodge", width = 0.75) +
  scale_fill_discrete(name = "", labels = c(expression(paste(pCO[2])), "pH")) + 
  labs(x="Month", y="Number of data (thousands)", title = "") +
  Mytheme(size_labs = 14,) +
  theme(axis.title.x = element_text(colour="black", size=14, face="bold"),
        axis.title.y = element_text(colour="black", size=14, face="bold")
  )
# Combine plots
p <- (p_avail / p_avail_month)
p <- p + plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size=14, face = "bold"))
ggsave(filename = "figures/availability.png", plot = p, width = 20, height = 15, units = "cm")
ggsave(filename = "figures/essd/availability.pdf", plot = p, width = 20, height = 15, units = "cm")

# deleted unneeded data
#rm(x, d_min, list=ls(pattern="^d2"))
```

>>>>

library("RColorBrewer")
if (!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
if (!require("seacarb")) install.packages("seacarb")
library("seacarb")
if (!require("gridExtra")) install.packages("gridExtra")
library("gridExtra")
if (!require("lubridate")) install.packages("lubridate")
library("lubridate")
if (!require("grid")) install.packages("grid")
library(grid)
if (!require("viridis")) install.packages("viridis")
library("viridis")
if (!require("dygraphs")) install.packages("dygraphs")
library("dygraphs")
if (!require("knitr")) install.packages("knitr")
require("knitr")
if (!require("lmodel2")) install.packages("lmodel2")
library("lmodel2")
if (!require("captioner")) install.packages("captioner")
library("captioner")
if (!require("xts")) install.packages("xts")
library("xts")
if (!require("scales")) install.packages("scales")
library("scales")
if (!require("skimr")) install.packages("skimr")
library("skimr")
if (!require("FNN")) install.packages("FNN")
library("FNN")
if (!require("patchwork")) install.packages("patchwork")
library("patchwork")
if (!require("pander")) install.packages("pander")
library("pander")
if (!require("tsibble")) install.packages("tsibble")
library("tsibble")
if (!require("slider")) install.packages("slider")
library("slider")
if (!require("kableExtra")) install.packages("kableExtra")
library("kableExtra")
if (!require("readxl")) install.packages("readxl")
library("readxl")
if (!require("oce")) install.packages("oce")
library(oce)
if (!require("pangaear")) install.packages("pangaear")
library("pangaear")
if (!require("latex2exp")) install.packages("latex2exp")
library("latex2exp")
if (!require("gtsummary")) install.packages("gtsummary")
library("gtsummary")
if (!require("todor")) install.packages("todor")
library("todor")
if (!require("ggpmisc")) install.packages("ggpmisc")
library("ggpmisc")
if (!require("tidyselect")) install.packages("tidyselect")
library("tidyselect")
if (!require("usethis")) install.packages("usethis")
library("usethis")
if (!require("ggOceanMaps")) install.packages("ggOceanMaps")
library("ggOceanMaps")
library(ggtext)  # remotes::install_github("clauswilke/ggtext")

@@

```{r, echo=FALSE, message= FALSE, warning=FALSE, out.width="80%",fig.show='hold', fig.align="center"}
# Continuous data first
d_min <- readRDS(file = paste0(path, "fb_awipev-co2_server/ny-alesund/data/NRT_data/all_nydata_minute.rds")) %>%
dplyr::select(datetime, c(pco2_corr_qf, sal_fb_qf, sal_insitu_ctd_qf, 
                          temp_fb_qf, temp_insitu_11m_qf, ph_dur_qf, phINT_qf))

s1 <- summary(as.factor(d_min$pco2_corr_qf))*100/length(d_min$pco2_corr_qf)
s2 <- summary(as.factor(d_min$sal_fb_qf))*100/length(d_min$sal_fb_qf)
s3 <- summary(as.factor(d_min$sal_insitu_ctd_qf))*100/length(d_min$sal_insitu_ctd_qf)
s4 <- summary(as.factor(d_min$temp_fb_qf))*100/length(d_min$temp_fb_qf)
s5 <- summary(as.factor(d_min$temp_insitu_11m_qf))*100/length(d_min$temp_insitu_11m_qf)
# durafet and seafet started later
x <- filter(d_min, datetime >= as.POSIXct("2017-08-29 10:00:00"))
#s6 <- summary(as.factor(x$temp_dur_qf))*100/length(x$temp_dur_qf)
s7 <- summary(as.factor(x$ph_dur_qf))*100/length(x$ph_dur_qf)
x <- filter(d_min, datetime >= as.POSIXct("2017-08-24 19:00:00"))
s8 <- summary(as.factor(x$phINT_qf))*100/length(x$phINT_qf)
# s7 not bound because durafet not in ESSD paper
z <- t(bind_rows(s1,s2,s3,s4,s5,s8)) %>%
  as.data.frame() %>%
         rownames_to_column(var = "flag") %>% 
         as_tibble()
labels <- c("Good", "Not usable", "Outside regional range", "Outliers (despike)", 
            "Failed visual inspection", "Not deployed or operated", "Acid flush", "Failed visual inspection")
colnames(z) <- c("Flag", "pCO2", "Salinity FerryBox", "In situ salinity", 
                 "Temperature FerryBox", "In situ temperature", "seaFET pH")
p_avail <- pivot_longer(z, -Flag) %>% 
  drop_na() %>% 
  ggplot(aes(x = name, y = value, fill = Flag)) +
  geom_bar(position = position_stack(reverse = TRUE), stat="identity", width = 0.3) +
  scale_x_discrete(
    labels = c(
      "Flag",
      'pCO2' = parse(text = latex2exp::TeX('$pCO_2$')),
      'In situ salinity' = "Salinity profiler",
      "Salinity Ferrybox",
      "Temperature FerryBox",
      'In situ temperature' = "Temperature 11 m",
      "seaFET pH" = "pH profiler"
    )
  ) +
  coord_flip() +
  labs(title = "A", x="", y = "Distribution of quality flags (%)") +
  Mytheme(size_labs = 14) +
  theme(legend.title = element_text(colour="black", size=14, face="bold"),
        axis.title.x = element_text(colour="black", size=14, face="bold"),
        title = element_text(colour="black", size=14, face="bold")
        )
d2 <- d_all %>% 
  dplyr::select(montht, pco2_fb, phINT)
count_pco2_fb_no_NA <- sum(!is.na(d2$pco2_fb)) # total non-NA pCO2 values
count_phINT_no_NA <-  sum(!is.na(d2$phINT)) # total non-NA pH seafet values
# calculate the distribution of non-NA pCO2 and pH values per month
d2_summary <- d2 %>% 
  dplyr::group_by(montht) %>% 
  dplyr::summarise(pco2 = round(100 * sum(!is.na(pco2_fb))/1000, digits = 1),
                   pH = round(100 * sum(!is.na(phINT))/1000, digits = 1),
                   pco2_pc = round(100 * sum(!is.na(pco2_fb))/count_pco2_fb_no_NA, digits = 1),
                   pH_pc = round(100 * sum(!is.na(phINT))/count_phINT_no_NA, digits = 1)
                   )
# plot the distribution of non-NA pCO2 and pH values per month
p_avail_month <- d2_summary %>% 
  tidyr::pivot_longer(cols = c(pco2, pH)) %>% 
  ggplot() +
  geom_col(aes(x = montht, y = value, fill = name), position = "dodge", width = 0.75) +
  scale_fill_discrete(name = "Continuous data", labels = c(expression(paste(pCO[2])), "pH")) + 
  scale_x_discrete(breaks = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"), 
                   labels = c("Jan", "", "Mar", "", "May", "", "Jul", "", "Sep", "", "Nov", "")) +
  labs(x="Month", y="Number of data (thousands)", title = "B") +
  Mytheme(size_labs = 14,) +
  theme(legend.title = element_text(colour="black", size=14, face="bold"),
        axis.title.x = element_text(colour="black", size=14, face="bold"),
        axis.title.y = element_text(colour="black", size=14, face="bold", margin = margin(r = -220)),
        title = element_text(colour="black", size=14, face="bold")
  )

# Discrete data now, only the monthly distribution
d3 <- d_all %>% 
  dplyr::select(montht, at, ct, ph_s_sf_temp_insi)

# calculate the distribution of non-NA values per month
d3_summary <- d3 %>% 
  dplyr::group_by(montht) %>% 
  dplyr::summarise(at = sum(!is.na(at)),
                   ct = sum(!is.na(ct)),
                   pH = sum(!is.na(ph_s_sf_temp_insi))
  )
# plot the distribution of non-NA pCO2 and pH values per month
p_discrete_avail_month <- d3_summary %>% 
  tidyr::pivot_longer(cols = c(at, ct, pH)) %>% 
  ggplot() +
  geom_col(aes(x = montht, y = value, fill = name), position = "dodge", width = 0.75) +
#  scale_fill_discrete(name = "Discrete data", labels = c(expression(paste(A[T]), C[T]), "pH")) + 
  scale_fill_discrete(name = "Discrete data", 
                      labels = c(TeX("\\textit{A}$_T$"), TeX("\\textit{C}$_T$"), "pH")) + 
  scale_x_discrete(breaks = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec"), 
                   labels = c("Jan", "", "Mar", "", "May", "", "Jul", "", "Sep", "", "Nov", "")) +
  labs(x="Month", y="Number of data", title = "C") +
  Mytheme(size_labs = 14,) +
  theme(legend.title = element_text(colour="black", size=14, face="bold"),
        axis.title.x = element_text(colour="black", size=14, face="bold"),
        axis.title.y = element_text(colour="black", size=14, face="bold", margin = margin(r = -220)),
        title = element_text(colour="black", size=14, face="bold")
  )

# Combine plots
p <- (p_avail / p_avail_month / p_discrete_avail_month)
#p <- p  #plot_annotation(tag_levels = 'A') & theme(plot.tag = element_text(size=14, face = "bold"))
ggsave(filename = "figures/availability.png", plot = p, width = 20, height = 15, units = "cm")
ggsave(filename = "figures/essd/availability.pdf", plot = p, width = 20, height = 20, units = "cm")

# deleted unneeded data
#rm(x, d_min, list=ls(pattern="^d2"))
```
Number of discrete data:

- AT: `r sum(!is.na(d_all$at))`
- CT: `sum(!is.na(d_all$ct))`
- Spectrophotometric pH: `sum(!is.na(d_all$ph_s_sf_temp_insi))`
